---
title: "Credit_Risk"
output: html_document
---

#install library
```{r install pkg,include=FALSE,cache=TRUE}
install.packages("dplyr", repos = "http://cran.us.r-project.org")
```

#####Credit Risk Modelling
Credit risk modelling is an important activity for financial institutions to help reduce the risk that a debtor will
default on a personal/commercial loan. One of the most widely used models in the credit scoring context is logistic
regression, which I will use in this project to explore and examine a sample credit data set.

library(readxl)
data <- read_excel("D:\\data science with R\\logistic regrssion\\loan_data.xlsx") 
str(data)  

#converting all char variables into factor in one go
```{r}
names <- c(3,5,7,9)
data[,names] <-lapply(data[,names], as.factor)
str(data)
'''
###------------------------------------ DATA PREPARATION STAGE ---------------------------------------------------- ### 

# Take a look at the structure of the data 
```{r}
str(data) 
```
# There are 11948 rows of 11 variables in this credit risk data set 

# Check the data set for any missing values 
```{r}
any(is.na(data)) # There is no missing data for this data set 
colSums(is.na(data))
summary(data)
```
#exploration
```{r}
v <-boxplot(data$annual_inc)
v$stats[1]
v$stats[5]

summary(data$annual_inc)
quantile(data$annual_inc)
quantile(data$annual_inc, probs = c(0.90, 0.95, 0.96, 0.97, 0.98))

data$annual_inc <- ifelse(data$annual_inc > 140000, 69333, data$annual_inc)
```
#still i find outlier at the top of boxplot
```{r}
data$annual_inc <- ifelse(data$annual_inc > 115000, 69333, data$annual_inc)

plot(data$annual_inc)
```
#we can see, annual_inc is positively skewed, means there are more no of people with low income, have to remove outliers


# Convert the term variable from a character to a categorical factor variable 
```{r}

levels(data$term)
length(data$term)

summary(data$term) 
data$term <- as.factor(data$term) 
```
# Convert the grade variable from a character to a factor variable representing 7 levels, 
# pertaining to a loan applicant's credit grade (A to G) representing best to worst credit grade 
```{r}

summary(data$grade)
data$grade <- as.factor(data$grade) 
table(data$grade) 
```
# Convert the home ownership variable from a character to a categorical factor variable 
```{r}
summary(data$home_ownership) 
data$home_ownership <- as.factor(data$home_ownership)
table(data$home_ownership)
```
# Convert the loan_status variable from a character to a factor variable 
```{r}

summary(data$loan_status)
data$loan_status <- as.factor(data$loan_status)
table(data$loan_status)
```

# To prepare the binary response variable for the logistic regression model, 
# create a new variable, charge_off, derived from the loan_status, wherein 1 denotes 
# a loan charge off (loan cannot be recovered by the bank) and 0 denotes FALSE (loan still outstanding) 
```{r}

library(dplyr)
data <- data %>% mutate(charge_off = ifelse(loan_status == "Charged Off", 1, 0))
```
# Examine the output of the newly created variable using a Cross Table
```{r}

library(descr)
CrossTable(data$charge_off, prop.r = T, prop.c = FALSE, prop.chisq = F, prop.t = F)
table(data$charge_off)
```

# The table shows that 12% of loans issued by the credit company defaulted and have been charged-off 
# There are 1428 cases of charge-off loans out of 11948 data points in the data set 
#proportions and cross table

```{r}

str(data)
table(data$charge_off)
summary(data$charge_off)

xtab1 <-xtabs(~charge_off + home_ownership, data)
xtab1
prop.table(xtab1)
```
#3rd variable
```{r}

xtab2 <-xtabs(~charge_off + home_ownership + grade, data)
xtab2
``` 
## flat table
```{r}

ftable(xtab2)
```
#SAS Like crosstable
```{r}

library(gmodels)
xtab1 <-xtabs(~charge_off + home_ownership, data)
CrossTable(xtab1)


y<-table(data$charge_off, data$home_ownership)
prop.table(y)

table(data$charge_off, data$home_ownership)
table(data$charge_off, data$grade)

table(data$charge_off, data$term)
x<-table(data$term, data$home_ownership)
prop.table(x,2) # 2 is for column
prop.table(x,1) # 1 is for row


plot(data$home_ownership,data$annual_inc)
plot(data$grade,data$annual_inc)
plot(data$charge_off,data$annual_inc)
```

#Data preparation
#have to create coding of categorized vraiables
#1 Term
```{r}

data <- data%>%mutate(threesixM = ifelse(term == "36 months", 1,0))
data <- data%>%mutate(sixzeroM = ifelse(term == "60 months", 1,0))

table(data$term)
table(data$threesixM)
table(data$sixzeroM)

```
#2home_ownership
```{r}

table(data$home_ownership)
data <- data%>%mutate(Hown = ifelse(home_ownership == "OWN", 1,0))
data <- data%>%mutate(Hrent = ifelse(home_ownership == "RENT", 1,0))
data <- data%>%mutate(Hmortg= ifelse(home_ownership == "MORTGAGE", 1,0))
table(data$Hown)

```


###------------------------------------ DATA SPLITTING --------------------------------------------------------------- ### 


# To make the training and test sets, set a seed using set.seed(). 
# set seed of 567 
```{r}

```{r}
set.seed(567) 
```
# Store row numbers for the TRAINING SET: index_train using 2/3 of the original data 
```{r}

index_train <- sample(1:nrow(data), 2/3 * nrow(data))
```

# Create the TRAINING SET as training_set containing 
```{r}

training_set <- data[index_train, ]
```
# Create the TEST SET as test_set 
```{r}

test_set <- data[-index_train, ]
```


###--------------------------------- LOGISTIC REGRESSION -------------------------------------------------------------- ### 


# Fit the Logistic Regression Model with the charge_off column as our binary response variable 
# Use the training_set to build the model 
```{r}

logistic.model <- glm(charge_off ~ home_ownership + annual_inc + loan_amnt + term + int_rate +
                        grade + fico_score + inq_last_6mths, family = "binomial", data = training_set)

str(training_set)
logistic.model <- glm(charge_off ~  annual_inc +fico_score + Hmortg + Hrent +threesixM + annual_inc + loan_amnt + int_rate +  inq_last_6mths, family = "binomial", data = training_set)
summary(logistic.model)

```
#annual_inc variable has p value 8.37e-12, less than 0.05, it means annual income appears to have signi impact
#on the probablity of debtors to have default

#lets check VIF and 
```{r}

lmodel <- lm(charge_off ~  annual_inc +fico_score + Hmortg + Hrent +threesixM + annual_inc + loan_amnt + int_rate +  inq_last_6mths, data = training_set)
library(car)
vif(lmodel)
```
#To test the hypothesis H0: b1= b2=0 
```{r}

anova(logistic.model, test="Chisq")
```
#As per output P value of Hrent variable isnt significant, so will remove this variable and will check

```{r}

logistic.model2 <- glm(charge_off ~  annual_inc +fico_score + Hmortg  +threesixM + annual_inc + loan_amnt + int_rate +  inq_last_6mths, family = "binomial", data = training_set)
summary(logistic.model2)

lmodel2 <- lm(charge_off ~  annual_inc +fico_score + Hmortg  +threesixM + annual_inc + loan_amnt + int_rate +  inq_last_6mths, data = training_set)
vif(lmodel2)

anova(logistic.model2, test="Chisq")
exp(coef(logistic.model2))

logistic.model3 <- glm(charge_off ~  annual_inc + threesixM + annual_inc  + int_rate +  inq_last_6mths, family = "binomial", data = training_set)
summary(logistic.model3)

```

###---------------------------- PREDICTING PROBABILITY OF A LOAN CHARGE-OFF ------------------------------------------ ### 


# Make predictions for the test set elements using the created logistic regression model
```{r}

predictions.logistic <- predict(logistic.model, newdata = test_set, type = "response") 
predictions.logistic2 <- predict(logistic.model2, newdata = test_set, type = "response") 
predictions.logistic3 <- predict(logistic.model3, newdata = test_set, type = "response") 
```


# Take a look at the range of the probability predictions 
```{r}

range(predictions.logistic)
range(predictions.logistic2)
range(predictions.logistic3)
```

# The range of predictions for both models is wide which is a good indicator; a small range means that the test set 
# cases do not lie far apart, therefore the model might not be good in discriminating good and bad loans 


###------------------------ EVALUATING THE RESULT OF LOGISTIC REGRESSION MODEL --------------------------------------- ### 

# To compare our predictions with the binary test_set$charge_off column, we must 
# transform the prediction vector to binary values of 1 and 0 indicating the status of the loan. 
# A cut-off or threshold must be set in this case. 
# If the predicted probability lies above the cutoff value then the prediction is set to 1, indicating
# a loan that charged off, otherwise it is set to 0, indicating the loan is still active;  
# A confusion matrix can be created afterwards to calculate Accuracy and compare cut-offs 
# The cut-off is basically a measure of risk tolerance of the financial institution 
# If we set a lower cutoff value, it means that we will classify a loan as a charge-off if it exceeds 
# a certain level of probabilistic risk. 

# USING A CUT-OFF VALUE OF 25% 
# Make a binary predictions-vector using a cut-off of 25%
```{r}

pred_cutoff_25 <- ifelse(predictions.logistic > 0.25, 1, 0)
pred_cutoff_25_1 <- ifelse(predictions.logistic2 > 0.25, 1, 0)
pred_cutoff_25_2 <- ifelse(predictions.logistic3 > 0.25, 1, 0)

```
# Construct a confusion matrix using a cut-off of 25% 
```{r}

conf_matrix_25 <- table(test_set$charge_off, pred_cutoff_25)
conf_matrix_25_1 <- table(test_set$charge_off, pred_cutoff_25_1)
conf_matrix_25_2 <- table(test_set$charge_off, pred_cutoff_25_2)

```
# Calculate for Accuracy 
```{r}

accuracy.25 <- sum(diag(conf_matrix_25)) / nrow(test_set)
accuracy.25_1 <- sum(diag(conf_matrix_25_1)) / nrow(test_set)
accuracy.25_2 <- sum(diag(conf_matrix_25_2)) / nrow(test_set)
```

# The accuracy for the model is 86.32% 

# USING A CUT-OFF VALUE OF 50% 
# Make a binary predictions vector using a cut-off of 50% 
```{r}

pred_cutoff_50 <- ifelse(predictions.logistic > 0.50, 1, 0) 
pred_cutoff_50_1 <- ifelse(predictions.logistic2 > 0.50, 1, 0) 
pred_cutoff_50_2 <- ifelse(predictions.logistic3 > 0.50, 1, 0) 
```

# Construct a confusion matrix using a cut-off of 50% 
```{r}

conf_matrix_50 <- table(test_set$charge_off, pred_cutoff_50)
conf_matrix_50_1 <- table(test_set$charge_off, pred_cutoff_50_1)
conf_matrix_50_2 <- table(test_set$charge_off, pred_cutoff_50_2)
```

# Calculate for Accuracy 
```{r}

accuracy.50 <- sum(diag(conf_matrix_50)) / nrow(test_set) 
accuracy.50_1 <- sum(diag(conf_matrix_50_1)) / nrow(test_set) 
accuracy.50_2 <- sum(diag(conf_matrix_50_2)) / nrow(test_set) 
```
# The accuracy for the model is 89.02% 
##